---
title: "Barbell Lift Accelerometer Analysis"
author: "Thomas Michem"
date: "26 December 2015"
output: html_document
---

<!-- TODO Publish on github -->
<!-- TODO a gh-pages branch so the HTML page can be viewed online -->
<!-- TODO Check < 2000 words -->
<!-- TODO Check <5 figures -->

# Description

Activity monitors can not only quantify  how much of an activity people do, but also how well they do it.

In this analysis the goal is to build a machine learning algorithm to predict activity quality from the activity monitors.


# Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).

# Data

The training data for this project are available here:

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

The test data are available here:

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>

The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.


```{r}
# getting the data
```


# Building the model

<!-- based on the "classe" variable in the training set. You may use any of the other variables to predict with. -->


# Cross validation

<!-- expect the out of sample error to be and estimate the error appropriately with cross-validation? -->

# Applying the model
<!--  TODO apply your machine learning algorithm to the 20 test cases available in the test data -->

# Submission Files

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```
