---
title: "Barbell Lift Accelerometer Analysis"
author: "Thomas Michem"
date: "26 December 2015"
output: html_document
---

<!-- TODO Check < 2000 words -->
<!-- TODO Check <5 figures -->

# Introduction - Project understanding

In this analysis the goal is to build a machine learning algorithm to predict activity quality from activity monitors.
Activity monitors can not only quantify  how much of an activity people do, but also how well they do it.

Groupware @ LES calls this Human Activity Recognition - HAR and has gathered data on the subject.
We will use this to build a model to assess the quality of barbell lift movements.

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Read more: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset)

# Data Understanding

The data is delivered from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

The data gathered by HAR is available in training and test csv's. 

```{r}
train.url <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
train <- read.csv(url(train.url), na.strings = c("NA", ""))

validate.url <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
validate <- read.csv(url(validate.url), na.strings = c("NA", ""))
````

The training data set contains the 'classe' variable which is a classification of the quality of execution of the barbell lift.

```{r}
table(train$classe)
```


# Data Preparation

Now the data is prepared for the modelling. This means:
- Selecting the attributes that can be used in the modelling
- Splitting the data in a training set and a cross validation testing set

## Attribute selection

Selecting the attributes for the model means two things:
1. Ignoring the attributes where there are values in few observations
2. Ignoring the identifying attributs (that are not related to movement)

### Attributes with few observations

For the model we ignore observations that have less than 10% of observations.

```{r}
obs.NA <- colMeans(is.na(train))
obs.NA.few <- obs.NA[obs.NA > 0.90]
length(obs.NA.few)
```

Of the 160 attributes, 100 have fewer than 10% observations.

```{r}
train <- train[,!(names(train) %in% names(obs.NA.few))]
validate <- validate[,!(names(validate) %in% names(obs.NA.few))]
dim(train)
dim(validate)
```

### Ignoring the identifiers

```{r}
train <- train[,7:60]
validate <- validate[,7:60]
```

### Training and testing set

The training data will be split into a training and testing data set.
```{r, message=FALSE}
library(caret)
```

```{r}
set.seed(1307)
train.partition <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
training <- train[train.partition,]
testing <- train[-train.partition,]
dim(training)
dim(testing)
```

# Modelling

The classe variable in the training set will be used to derive a model that predicts the values. 
This is a typical problem for a random forest model.

```{r, echo=FALSE, message=FALSE}
set.seed(1983)
library(randomForest)
```

First we build the model.
```{r}
model.RF <- randomForest(classe~., data=training)
```

# Evaluation

## Cross validaton
Then we use the model to predict the values of the testing data set. The outcome is visualised in a confusion matrix.
```{r}
predict.RF <- predict(model.RF, newdata=testing)
confusionMatrix(predict.RF  , testing$classe)
```

The model has an accuracy of 0.9966. The accuracy is high for all classes.

## Applying the model

```{r}
predict.RF.final <- predict(model.RF, newdata=validate)
```

# Deployment

Deployment for this project is creating submission files in which the 20 test cases are validated against the model.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predict.RF.final)
```
